The response theory for the system under a non-equilibrium driving

### 2013-06-26 writedown/linear.response/paper.tex
add finite diff SD

\section{Numerical tests: the core set identification of  the Markov State Model }

The core set plays a very important role in the MSM theory.  We
therefore want to test our idea of non-equilibrium response theory by
identifying the core sets in a non-equilibrium driven system.
There are several ways to define a core set.
One possible choice  is to calculate the stability of the phase space
distribution: $\Delta \mt_\tau f(\vect x, t) = (\mt_\tau -\mi) f(\vect
x, t)$. And the core sets are defined by
\begin{align}
  C_\tau(t) = \{\vect x\vert \Delta \mt_\tau f(\vect x, t) \geq 0\}
\end{align}
Here the propagator $\mt_\tau$ is defined by $\mt_\tau =
e^{\tau\fwg_t^{2\beta}}$.  Notice that here the physical time $t$ is
frozen, and the distribution $f$ evolves a long the artificial time
$\tau$ at temperature $1/(2\beta)$. The artificial time $\tau$ should
be chosen smaller than the investigated relaxation time scales of the
original system. However, on the other hand, $\tau$ cannot be
arbitrarily small, because otherwise the statistical uncertainty would
not allow us to distinguish $\mt_\tau f$ and $f$.

\subsection{The tilting double-well potential}

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{figs/fig-tilt-pot.eps}
  \caption{The double-well potential with tilting perturbation at
    time $t\geq 20$ ps..}
  \label{fig:tmp1}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/fig-tilt-str2-2d-more.eps}
  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$ under
    perturbation with strength $\eps = 2$.  From left to right the
    columns present results at time $t = 0$, 10, 15 and
    20~\textsf{ps}.  The 1st row: the direct non-equilibrium
    simulation. The 2nd row: the results of traditional equilibrium
    linear response theory.  The 3rd row: the results of the response
    theory developed by this work.  We calculate response terms up to
    the second order from an equilibrium simulation (i.e. letting $k_e
    = 0$ and $\eps = 0$ in Eq.~\eqref{eqn:langevin-2b}).
  }
  \label{fig:tmp2}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/fig-tilt-str4-2d-more.eps}
  \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
    $\eps = 4$. From left to right the
    columns present results at time $t = 0$, 10, 15 and
    20~\textsf{ps}.
    The first three rows use the same algorithm as Fig.~\ref{fig:tmp2}.    
    The 4th row: the results of the response theory developed by this work.
    Instead of calculating the responses from the equilibrium simulation,
    here we calculate the linear response from a non-equilibrium simulation
    of strength 2 (i.e. $k_e = 2$ and $\eps = 4$
    in Eq.~\eqref{eqn:langevin-2b}).
    Therefore, the effective perturbation in this case is only
    $\eps = 2$.
  }
  \label{fig:tmp3}
\end{figure}

We test the idea firstly by a one-dimensional model system: one particle in a
tilting double-well potential. For convenience, we let the mass of the
particle to be 1~\textsf{amu},
and the friction coefficient to be $1\,\textsf{ps}^{-1}$.
The unperturbed
Hamiltonian of the system is given by:
\begin{align}
  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
\end{align}
with potential
\begin{align}
  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
\end{align}
Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
At room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
The red line in Fig.~\ref{fig:tmp1} presents the unperturbed potential.
%$\textsf{kJ} / (\textsf{mol nm}^4)}$
The perturbation is given by
\begin{align}
  \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) = 1
\end{align}
Here $V(\vect q) = -\vect q$   effectively tilts the original
potential $U(\vect q)$ (see the gree and blue lines in Fig.~\ref{fig:tmp1}).
The strength of $F_e(t)$ is
set to be $k_e$, when the time $t$ is larger than some warm-up time $t_c$.
When $t \leq t_c$, the strength of $F_e(t)$ is linearly growing,
i.e. $ F_e(t) = k_e\cdot t/t_c$.
We use the value $t_c = 20$~\textsf{ps}, so that the perturbation increases slow
enough: the typical decaying time scale of the correlation function
in~\eqref{eqn:eq-lr} is roughly 3~\textsf{ps}.
We choose $\tau = 1$~\textsf{ps}, and
% We choose $\mt_\alpha$
% to be the propagator of the Langevin dynamics at temperature $150$
% \textsf{K} with $\alpha = 1$ \textsf{ps}.
consider the stability of the
distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
The larger
this value, the more stable the core sets are.

See Fig.~\ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
the results of the brute-force non-equilibrium simulation
and that of the response theory are presented.
In Fig.~\ref{fig:tmp2},
the good agreement between the brute-force (1st row)
and the linear response result (2nd row) indicates that $\eps = 2$ is
indeed a relatively small perturbation to the equilibrium.
In the third row of Fig.~\ref{fig:tmp2},
the second order response is considered. We do not see any
substantial improvement of the result by including this higher order term,
because the
driving force to the system is small,
and the linear term is anyway a good approximation.

In Fig.~\ref{fig:tmp3}, as the
perturbation grows stronger, the left well becomes less stable, while
the right well becomes more stable.  When the perturbation is as strong as
$\eps = 4$, the linear response result (2nd row) is no
longer precise:
At $t\geq 20$ \textsf{ps}, the left well basically
disappears, however, the linear response calculation presents
artificial structure (that is a small artificial
stable region accompanied by a small unstable region)
at the position of the left well.
The second order response (3rd row) improves the accuracy a little, 
however, it also includes artificial finer-grained oscillation in the profile.
At the same time, the numerical uncertainty is higher, so the figure
looks noisy.
In the 4th row of Fig.~\ref{fig:tmp3}, we present the linear response result
that calculate from 
a intermediate
non-equilibrium simulation with driving force of $k_e = 2$.
Notice that the reference simulation in this case is also a
non-equilibrium simulation, so the effective strength of perturbation
is reduced to $\eps = 2$. According to Fig.~\ref{fig:tmp1}, this can be
considered as a small perturbation.
A substantial improvement is achieved: the profile
is correctly calculated, and the statistical error is small.

% the 1st order result of the
% response formula~\eqref{eqn:pert-approx-1} and
% \eqref{eqn:pert-approx-2}, with
% respect to the reference state $F_e^{\textrm{max}} = 2$. The result
% is impressively improved compared with the linear response with respect to the
% equilibrium state.

\subsection{Splitting single-well potential}

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
  \caption{The single-well potential with splitting perturbation at
    time $t\geq 20$ ps.}
  \label{fig:tmp4}
\end{figure}

% \begin{figure}
%   \centering
%   \includegraphics[width=0.95\textwidth]
%   {figs/fig-split-str1-2d-more.eps}
%   \caption{The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
%     $F_e^{\textrm{max}} = 1.0$.
%     The 1st row: the direct non-equilibrium
%     simulation. The 2nd row: the results of linear response
%     formula~\eqref{eqn:core-identify-approx-2}.
%     The 3rd row: the results of response
%     formula~\eqref{eqn:pert-approx-1} and \eqref{eqn:pert-approx-2}.
%     The order is 2, the reference state is $F_e^{\textrm{max}} = 0$, i.e.
%     the equilibrium state.
%   }
%   \label{fig:tmp5}
% \end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/fig-split-str2-2d-more.eps}
  \caption{
    The plot of $\Delta\mt_\tau f(\vect x,t)$  under perturbation of
    $\eps = 2$. From left to right the
    columns present results at time $t = 0$, 5, 10 and
    20~\textsf{ps}.
    The 1st row: the direct non-equilibrium
    simulation. The 2nd row: the results of traditional equilibrium
    linear response theory.  The 3rd row: the results of the response
    theory developed by this work.  We calculate response terms up to
    the second order from an equilibrium simulation (i.e. letting $k_e
    = 0$ and $\eps = 2$ in Eq.~\eqref{eqn:langevin-2b}).
    The 4th row: the results of the response theory developed by this work.
    Instead of calculating the responses from the equilibrium simulation,
    here we calculate the linear response from a non-equilibrium simulation
    of strength 2 (i.e. $k_e = 1$ and $\eps = 1$
    in Eq.~\eqref{eqn:langevin-2b}).
    Only the linear term of response is calculated.
    Notice, the effective perturbation in this case is only
    $\eps = 1$.
  }
  \label{fig:tmp6}
\end{figure}

In the previous subsection, we considered an example that the stability of the
core sets is switched by an external driving to the system.
In this section, we consider ``splitting single-well potential'' example
that two new core sets grows out of 
an old core sets by the external driving force.
In this example, the unperturbed
potential is simply the harmonic potential given by:
\begin{align}
  U(\vect q) = \frac12\,k\,\vect q^2 
\end{align}
Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
See the red line in Fig.~\ref{fig:tmp4} for the unperturbed potential.
%$\textsf{kJ} / (\textsf{mol nm}^4)}$
The perturbation has again given by
\begin{align}
  \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) 
\end{align}
where $V(\vect q)$ has a Gaussian profile:
\begin{align}
  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
\end{align}
we use $\sigma = 0.16$~\textsf{nm}.

In this example, we also consider two different strengths of
perturbations. Please see the green and blue lines in Fig~\ref{fig:tmp4} for
perturbations of $\eps = 1$ and $\eps = 2$, respectively.
Numerical results show that the perturbation $\eps=1$ can be considered to be
relatively small. Since the phenomenon is basically the same as
Fig.~\ref{fig:tmp2}, we do not present them here.
Fig.~\ref{fig:tmp6} presents the numerical results of perturbation $\eps =
2$.  Only with a linear response calculated from the equilibrium
simulation (2nd row), the stability of the two new core sets are too
strong, and the unstable region between them is over estimated, comparing with
the brute force simulation presented on the first row.
We present the  second order response results on the 3rd row, which is
also calculated from the equilibrium simulation. In this case, the statistical
error is too large, so that
the profile of $\mt_\tau f$ is indeed too noisy to draw any useful information.
On the 4th row of Fig.~\ref{fig:tmp6} presents the linear response starting
from a non-equilibrium simulation with external driving $k_e = 1$.
In this case the effective strength of the perturbation to the reference
simulation is $\eps = 1$, which is considered to be relatively small.
The numerical results is satisfactorily consistent with the brute force
non-equilibrium simulation.

\subsection{Optimal forcing}
\begin{figure}
  \centering
  \includegraphics[]{figs/fig-ctr-stat.eps}
  \caption{The optimal forcing.  In this plot, $T=20$ ps.  The red
    color indicates $\eta = 0.01$ and the green color indicates $\eta
    = 0.05$.  In the upper plot, the solid lines present the magnitude
    of the control $F_e(t)$, which is calculated on a family of
    piecewise linear ansatz functions. The time interval of a linear
    piece is 1 ps.  In the lower plot, the dashed lines display the
    value of $P(\vect X_t)$ as the time evolves.  The solid lines
    present $O[F_e](t)$.  }\label{fig:tmp7}
\end{figure}

We investigate the example of the double well potential, and
consider the problem of optimally forcing the population from the
left well to the right well. That is equal to consider the 
the following optimal forcing problem:
\begin{align}
  O_{\max}(T)& = \max_{F_e\in\mathcal F} O[F_e](T) 
\end{align}
with
\begin{align}
  O[F_e](T) &= P(\vect X_T) - \eta \int_0^T\vert F_e(t)\vect D(\vect q)\vert^2\, dt
\end{align}
where $P(\vect X_T)$ is equal to $1$ when the end point of the
trajectory $\vect X_T$ is in the right well, and equal to 0 otherwise.
It is clear that for an unbiased double-well, the average of $P(\vect
X_T)$ is 0.5.  The integral is the ``cost'' of the control and $\eta$
is a positive constant that indicates the relative magnitude of the
cost. We consider the ansatz function family $\mathcal F$ of piecewise
linear functions on $[0, T]$.

>>>>

\section{Numerical tests}

\subsection{The metastability of a single-well potential}
The metastability of the system at time $t$ can be investigated by
calculating the function $\Delta \mt_\tau f(\vect x, t) = (\mt_\tau
-\mi) f(\vect x, t)$, where $f(\vect x, t)$ is the probability
distribution at position $\vect x$ and time $t$, the propagator
$\mt_\tau$ is defined by $\mt_\tau = e^{\tau\fwg_t^{2\beta}}$, and
$\mi$ is the identity.  Notice that here the physical time $t$ is
frozen, and the distribution $f$ evolves a long the artificial time
$\tau$ at temperature $1/(2\beta)$.  If the position $\vect x$ is
``stable'', then intuitively the probability of the system being at
position $\vect x$ will be higher when the temperature is lower, i.e.,
$\Delta \mt_\tau f(\vect x, t) > 0$.  Otherwise, when position $\vect
x$ is not stable, $\Delta \mt_\tau f(\vect x, t) < 0$.  The
artificial time $\tau$ should be chosen smaller than the investigated
relaxation time scales of the original system. However, on the other
hand, $\tau$ cannot be arbitrarily small, because otherwise the
statistical uncertainty would not allow us to distinguish $\mt_\tau f$
and $f$.

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{figs/fig-split-pot.eps}
  \caption{The single-well potential with splitting splitting drving
    force at time $t_c = 20$ ps.  Since the drving force is
    conservative ($\vect D(\vect q) = -\nabla_{\vect q}V(\vect q)$),
    we plot the non-equilibrium drving energy of the system.  The red,
    green and blue lines are the potential energy without
    non-equilibrium driving, with non-equilibrium driving and
    the non-equilibrium driving perturbed by $\eps\Delta F_e(t)$,
    respectively.  }
  \label{fig:tmp4}
\end{figure}

% The core set plays a very important role in the MSM theory.  We
% therefore want to test our idea of non-equilibrium response theory by
% identifying the core sets in a non-equilibrium driven system.
% There are several ways to define a core set.
% One possible choice  is to calculate the stability of the phase space
% distribution: $\Delta \mt_\tau f(\vect x, t) = (\mt_\tau -\mi) f(\vect
% x, t)$. And the core sets are defined by
% \begin{align}
%   C_\tau(t) = \{\vect x\vert \Delta \mt_\tau f(\vect x, t) \geq 0\}
% \end{align}
% Here the propagator $\mt_\tau$ is defined by $\mt_\tau =
% e^{\tau\fwg_t^{2\beta}}$.  Notice that here the physical time $t$ is
% frozen, and the distribution $f$ evolves a long the artificial time
% $\tau$ at temperature $1/(2\beta)$. The artificial time $\tau$ should
% be chosen smaller than the investigated relaxation time scales of the
% original system. However, on the other hand, $\tau$ cannot be
% arbitrarily small, because otherwise the statistical uncertainty would
% not allow us to distinguish $\mt_\tau f$ and $f$.

We use the idea of non-equilibrium linear response to investigate the
metastability a one-dimensional model system: one particle in a
splitting single-well potential.  For convenience, we let the mass of
the particle to be 1~\textsf{amu}, and the friction coefficient to be
$1\,\textsf{ps}^{-1}$.
The temperature is the
room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
The unperturbed Hamiltonian of the system is
given by:
\begin{align}
  \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
\end{align}
with potential
\begin{align}
  U(\vect q) = \frac12\,k\,\vect q^2 
\end{align}
Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^2)$.
See the red line in Fig.~\ref{fig:tmp4} for the  potential $U$.
%$\textsf{kJ} / (\textsf{mol nm}^4)}$
The non-equilibrium driving $\vect D$ is given by a conservative force:
\begin{align}
  \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) 
\end{align}
where the driving potential $V(\vect q)$ has a Gaussian profile:
\begin{align}
  V(\vect q) = \frac{1}{\sqrt{2\pi \sigma^2}}
  \exp\Big\{-\frac{\vect q^2}{2\sigma^2}\Big\}
\end{align}
we use $\sigma = 0.16$~\textsf{nm}.  The strength of non-equilibrium
driving $F_e(t)$ is set to be a unitless constant $k_e$, when the time $t$ is
larger than some warm-up time $t_c$.  When $t \leq t_c$, the strength
of $F_e(t)$ is linearly growing, i.e. $ F_e(t) = k_e\cdot t/t_c$.  We
consider the perturbation to the system given by $\eps\Delta F_e(t) =
\eps$ when $t> t_c$, and $\eps\Delta F_e(t) = \eps \cdot t/t_c$ when
$t\leq t_c$.
We consider the following parameters: $t_c = 20$~\textsf{ps}, 
$k_e = 1$ and $\eps = 1$, see Fig.~\ref{fig:tmp4}
for the non-equilibrium driving potential and perturbed
potential at time $t = t_c$.
$\tau$ is set to be $1~\textsf{ps}$.

\begin{figure}
  \centering
  \includegraphics[width=0.95\textwidth]
  {figs/fig-split-str2-2d-more.eps}
  \caption{ The plot of $\Delta\mt_\tau f(\vect x,t)$ in the phase
    space under perturbed non-equilibrium driving ($k_e = 1$ and $\eps
    = 1$). From left to right the columns present results at time $t =
    0$, 5, 10 and 20~\textsf{ps}.  The 1st row: the brute force
    non-equilibrium simulation. The 2nd row: the results of
    traditional equilibrium linear response theory. In this case we
    effectivly consider an equilibrium system free of driving ($k_e =
    0$), and the perturbation $\eps$ is set to 2.  The 3rd row: the
    second order equilibrium response result.  The 4th row: the
    non-equilibrium linear response result.  }
  \label{fig:tmp6}
\end{figure}

Fig.~\ref{fig:tmp6} presents the numerical results of the splitting
single-well potential. From left to right the four rows presents the
stability of the system at time $t = 0$, 5, 10 and 20~\textsf{ps}. The first
row presents the brute force non-equilibrium simulation.  It is clear
that at the beginning, there is only one stable region in the phase,
i.e. arount $q = 0$. As the time evolves, an energy barrier develops
in the center of the simulation region, therefore, the single stable
region splits into two equally sized stable regions.  In the end, the two
stable regions are entirly seperated.  This brute non-equilibrium
simulation serves as the precise result that the response theory
should compare with. The second row shows the result of the
traditional equilibrium linear response theory.  Please notice that in
this case, since the reference simulation is in equilibrium (i.e. $k_e = 0$),
the effective perturbation is of strength $\eps = 2$.  At $t =
5$~\textsf{ps}, the accuracy of the equilibrium linear response is
perfect. When $t = 10$~\textsf{ps}, the left stable region is a little
over estimate. This may be due to the statistical uncertainty. At $t =
20$~\textsf{ps}, stability of the two stable regions are too strong,
and the unstable region in between is less stable than the brute
force result. The accuracy of the equilibrium linear response is poor
at $t \geq 10$~\textsf{ps}, because the strength of perturbation is no
longer small so that the preliminary assumption of the reponse theory
is not well fulfilled. On the third row, we present second order
equilibrium response theory. The using of a higher order response term
should improve the numerical accuracy, however, this is not the case,
because the second order response term cannot be accuratly calculated
due to the huge statistical error. By doing longer reference
equilibrium simulations may improve the accuracy of statistics, but it
is not ecnomic to do that.  On the fourth row of Fig.~\ref{fig:tmp6},
we present the linear response starting from a non-equilibrium driving
of $k_e = 1$.  The numerical results is satisfactorily consistent with
the brute force non-equilibrium simulation, because the perturbation
($\eps = 1$) is so small that the linear perturbation theory achieves
good accuracy.

% Fig.~\ref{fig:tmp6} presents the numerical results of perturbation $\eps =
% 2$.  Only with a linear response calculated from the equilibrium
% simulation (2nd row), the stability of the two new core sets are too
% strong, and the unstable region between them is over estimated, comparing with
% the brute force simulation presented on the first row.
% We present the  second order response results on the 3rd row, which is
% also calculated from the equilibrium simulation. In this case, the statistical
% error is too large, so that
% the profile of $\mt_\tau f$ is indeed too noisy to draw any useful information.

% \subsection{The tilting double-well potential}

% We test the idea firstly by a one-dimensional model system: one particle in a
% tilting double-well potential. For convenience, we let the mass of the
% particle to be 1~\textsf{amu},
% and the friction coefficient to be $1\,\textsf{ps}^{-1}$.
% The unperturbed
% Hamiltonian of the system is given by:
% \begin{align}
%   \mh (\vect p, \vect q) = \frac 12 \vect p^2 + U(\vect q) 
% \end{align}
% with potential
% \begin{align}
%   U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
% \end{align}
% Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\ \textsf{nm}$.
% At room temperature $300\ \textsf{K}$, $k_BT = 2.48$~\textsf{kJ/mol}.
% The red line in Fig.~\ref{fig:tmp1} presents the unperturbed potential.
% %$\textsf{kJ} / (\textsf{mol nm}^4)}$
% The perturbation is given by
% \begin{align}
%   \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) = 1
% \end{align}
% Here $V(\vect q) = -\vect q$   effectively tilts the original
% potential $U(\vect q)$ (see the gree and blue lines in Fig.~\ref{fig:tmp1}).
% The strength of $F_e(t)$ is
% set to be $k_e$, when the time $t$ is larger than some warm-up time $t_c$.
% When $t \leq t_c$, the strength of $F_e(t)$ is linearly growing,
% i.e. $ F_e(t) = k_e\cdot t/t_c$.
% We use the value $t_c = 20$~\textsf{ps}, so that the perturbation increases slow
% enough: the typical decaying time scale of the correlation function
% in~\eqref{eqn:eq-lr} is roughly 3~\textsf{ps}.
% We choose $\tau = 1$~\textsf{ps}, and
% % We choose $\mt_\alpha$
% % to be the propagator of the Langevin dynamics at temperature $150$
% % \textsf{K} with $\alpha = 1$ \textsf{ps}.
% consider the stability of the
% distribution $f(\vect x, t)$ at time $t$: $\Delta\mt_\tau f(\vect x, t)$.
% The larger
% this value, the more stable the core sets are.

% See Fig.~\ref{fig:tmp2} and \ref{fig:tmp3} for numerical result.  Both
% the results of the brute-force non-equilibrium simulation
% and that of the response theory are presented.
% In Fig.~\ref{fig:tmp2},
% the good agreement between the brute-force (1st row)
% and the linear response result (2nd row) indicates that $\eps = 2$ is
% indeed a relatively small perturbation to the equilibrium.
% In the third row of Fig.~\ref{fig:tmp2},
% the second order response is considered. We do not see any
% substantial improvement of the result by including this higher order term,
% because the
% driving force to the system is small,
% and the linear term is anyway a good approximation.

% In Fig.~\ref{fig:tmp3}, as the
% perturbation grows stronger, the left well becomes less stable, while
% the right well becomes more stable.  When the perturbation is as strong as
% $\eps = 4$, the linear response result (2nd row) is no
% longer precise:
% At $t\geq 20$ \textsf{ps}, the left well basically
% disappears, however, the linear response calculation presents
% artificial structure (that is a small artificial
% stable region accompanied by a small unstable region)
% at the position of the left well.
% The second order response (3rd row) improves the accuracy a little, 
% however, it also includes artificial finer-grained oscillation in the profile.
% At the same time, the numerical uncertainty is higher, so the figure
% looks noisy.
% In the 4th row of Fig.~\ref{fig:tmp3}, we present the linear response result
% that calculate from 
% a intermediate
% non-equilibrium simulation with driving force of $k_e = 2$.
% Notice that the reference simulation in this case is also a
% non-equilibrium simulation, so the effective strength of perturbation
% is reduced to $\eps = 2$. According to Fig.~\ref{fig:tmp1}, this can be
% considered as a small perturbation.
% A substantial improvement is achieved: the profile
% is correctly calculated, and the statistical error is small.

% % the 1st order result of the
% % response formula~\eqref{eqn:pert-approx-1} and
% % \eqref{eqn:pert-approx-2}, with
% % respect to the reference state $F_e^{\textrm{max}} = 2$. The result
% % is impressively improved compared with the linear response with respect to the
% % equilibrium state.

\subsection{Optimal tilting of a double-well potential}
\begin{figure}
  \centering
  \includegraphics[]{figs/fig-ctr-stat.eps}
  \caption{The optimal forcing.  In this plot, $T=20$ ps.  The red
    color indicates $\eta = 0.01$ and the green color indicates $\eta
    = 0.05$.  In the upper plot, the solid lines present the magnitude
    of the control $F_e(t)$, which is calculated on a family of
    piecewise linear ansatz functions. The time interval of a linear
    piece is 1 ps.  The blue insertions show the shape of the
    non-equilibrium driving potential $U(\vect q) + F_e(t)V(\vect
    q)$ at time indicated by the black arrow.  In the lower plot, the
    dashed lines display the value of $P(\vect X_t)$ as the time
    evolves.  The solid lines present $O[F_e](t)$.  }\label{fig:tmp7}
\end{figure}

In this section, we consider the following double well potential:
\begin{align}
  U(\vect q) = \frac12 k (\vect q^2 - a^2)^2
\end{align}
Here $k = 8$~$\textsf{kJ} / (\textsf{mol nm}^4)$, and $ a = 1\
\textsf{nm}$.  See the leftmost blue insertion of Fig.~\ref{fig:tmp7} for
the shape of the potential.
%$\textsf{kJ} / (\textsf{mol nm}^4)}$
The perturbation is given by a conservative tilting to $U(\vect q)$:
\begin{align}
  \vect D(\vect q) = -\nabla_{\vect q} V(\vect q) = 1
\end{align}
Here $V(\vect q) = -\vect q$. We want to
optimally design the tilting such that the probability
of being in the right well is as high as possible 
at the end time of the process. That is equal to consider the 
the following optimal forcing problem:
\begin{align}
  O_{\max}(T)& = \max_{F_e\in\mathcal F} O[F_e](T) 
\end{align}
with
\begin{align}
  O[F_e](T) &= P(\vect X_T) -
  \eta\,
  \big\langle \int_0^T\vert F_e(t)\vect D(\vect q)\vert^2\, dt\,
  \big\rangle
\end{align}
where $P(\vect X_T)$ is probability of the end point of the trajectory
$\vect X_T$ being in the right well.  It is
clear that for an unbiased double-well, $P(\vect X_T)$
is 0.5.  The integral is the ``cost'' of the control and $\eta$ is a
positive constant that indicates the relative magnitude of the
cost. We consider the ansatz function family $\mathcal F$ of piecewise
linear functions on $[0, T]$.
